---
title: "Scatter plot in R using Plotly"
author: "Wei Fan"
date: "June 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an R tutorial to help beginners to make scatter plots using the **Plotly** Package.

## Set up

### 0. Install Plotly Package adn add library

First, let's install the "Plotly" Package. If you have already installed the Plotly package you can skip the "install" step. Since we only want to install a package once, we can type the following commad in the R **Console**, instead of the R script.

```{r eval=FALSE}
install.packages("plotly")
```

After install the package, we can add the Plotly library. Libraries are usually added at the begging of the R scripts.

```{r message=FALSE}
library(plotly)
```

### 1. Loading dataset

For this tutorial, we will be using a buiilt-in R data: _mtcars_. Let's load the data, and have a closer look to out dataset. If you want to see more lines from the dataset, you can replace the number _3_ in the **head** function with the number of lines you want to see, or you can use the **View** function. 

```{r}
# Load dataset
data("mtcars")
# Examine the first 3 rows of the data
head(mtcars, 3)

```

If you want to see the full dataset, type:

```{r eval=FALSE}
View(mtcars)
```

We can get the information of this dataset using the code:

```{r eval=FALSE}
?mtcars
```

The **mtcars** dataset contains 32 observations and 11 features:
1. mpg	 Miles/(US) gallon
2. cyl	 Number of cylinders
3. disp	 Displacement (cu.in.)
4. hp	 Gross horsepower
5. drat	 Rear axle ratio
6. wt	 Weight (1000 lbs)
7. qsec	 1/4 mile time
8. vs	 V/S
9. am	 Transmission (0 = automatic, 1 = manual)
10. gear	 Number of forward gears
11. carb	 Number of carburetors

## Draw a Scatter Plot in R Using PLotly 

In this example we will plot a scatter plot of miles per gallon versus displacement. Below, we set column _mpg_ as x axis and column _disp_ as y axis. We set the dots in the scatter of size _10_ and color _red_. In the **layout()**, we set plot title and names of the x axis and y axis. 

```{r}
plot_ly(data = mtcars, x = ~mpg, y = ~disp, type = "scatter", mode = "markers", marker = list(size = 10, color = "red")) %>% 
  layout(title = "Miles per gallon versus displacement", xaxis = list(title = "Miles per gallon"), 
         yaxis = list(title = "Displacement"))

```

We can also add a **colorbar** _"Miles per gallon"_ to the scatter plot and let the colors of the dots changed by the value of column _cyl_. We can set the **colorscale** of the colorbar as _'Viridis'_. 

```{r}
plot_ly(data = mtcars, x = ~mpg, y = ~disp, type = "scatter", mode = "markers", marker = list(size = 10, colorbar = list(title = "Number of cylinders"), color = ~cyl, colorscale='Viridis', reversescale =T)) %>% 
  layout(title = "Miles per gallon versus displacement", xaxis = list(title = "Miles per gallon"), 
         yaxis = list(title = "Displacement"))

```

Similarly we can let the **size** of dots depends on the weigt, the _gear_ column.

```{r}
plot_ly(data = mtcars, x = ~mpg, y = ~disp, type = "scatter", mode = "markers", marker = list(size = ~gear, colorbar = list(title = "Number of cylinders"), color = ~cyl, colorscale='Viridis', reversescale =T)) %>% 
  layout(title = "Miles per gallon versus displacement", xaxis = list(title = "Miles per gallon"), 
         yaxis = list(title = "Displacement"))

```

Then, we realizes these dots are too small, we can increase the size of the dots inscale by add "*4":
```{r}
plot_ly(data = mtcars, x = ~mpg, y = ~disp, type = "scatter", mode = "markers", marker = list(size = ~gear*4, colorbar = list(title = "Number of cylinders"), color = ~cyl, colorscale='Viridis', reversescale =T)) %>% 
  layout(title = "Miles per gallon versus displacement", xaxis = list(title = "Miles per gallon"), 
         yaxis = list(title = "Displacement"))

```